---
title: "Continuous Distributions with R"
author: "Tripp Bishop/STAT 5255"
fig-align: center
format: pdf
execute: 
  warning: false
  message: false
---

```{r initial setup}
library(tidyverse)
library(latex2exp)
rm(list = ls())
theme_set(theme_minimal())
```

# Problem 6.27
Sketch the graphs of the beta densities having  
a) $\alpha=2$, $\beta=2$
b) $\alpha=1/2$, $\beta=1$
c) $\alpha=2$, $\beta=1/2$
d) $\alpha=2$, $\beta=5$

```{r problem 6.27}
X <- seq(from=0,to=1,by=0.01)

Y.a <- dbeta(X, 2, 2)
Y.b <- dbeta(X, 1/2, 1)
Y.c <- dbeta(X, 2, 1/2)
Y.d <- dbeta(X, 2, 5)

tibble(X,Y.a,Y.b,Y.c,Y.d) |> 
  ggplot(aes(x=X)) +
    geom_line(aes(y=Y.a), colour="magenta", linewidth=1) + 
    geom_line(aes(y=Y.b), colour="steelblue", linewidth=1) + 
    geom_line(aes(y=Y.c), colour="orange", linewidth=1) + 
    geom_line(aes(y=Y.d), colour="seagreen", linewidth=1) +
    annotate(geom="text", label=TeX("$\\alpha=1/2, \\beta=1$"), x=0.09, y=6, colour="steelblue") +
    annotate(geom="text", label=TeX("$\\alpha=2, \\beta=1/2$"), x=0.885, y=5, colour="orange") +
    annotate(geom="text", label=TeX("$\\alpha=2, \\beta=5$"), x=0.19, y=2.75, colour="seagreen") +
    annotate(geom="text", label=TeX("$\\alpha=2, \\beta=2$"), x=0.565, y=1.79, colour="magenta") +
    labs(
      title = TeX("$\\beta$ densities"),
      y = "Density"
    )
```
# Problem 6.52
In a certain city, the daily consumption of electric power in millions of 
kilowatt-hours can be treated as a random variable having a gamma distribution
with $\alpha=3$ and $\beta=2$. If the power plant of this city has a daily 
capacity of 12 million kilowatt-hours, what is the probability that this power
supply will be inadequate on any given day?

What is the probability that $X > 12 MWh/day$ given $X\sim \Gamma(3,2)$?

```{r problem 6.52}
(prob <- pgamma(12, shape=3, scale=2, lower.tail = FALSE))
```
The probability that power demand exceeds the power plant's capacity on any 
given day is $p=`r round(prob, 4)`$.

# Problem 6.54
The amount of time that a watch will run without having to be reset is a random
variable having an exponential distribution with $\theta=120$ days. Find the 
probabilities that such a watch will  
a) have to be reset in less than 24 days.

```{r problem 6.54a}
(prob <- pexp(24, rate = 1/120))
```
The probability that the watch will have to be reset in less than 24 days is
$p=`r round(prob,4)`$.

b) not have to be reset in at least 180 days.

```{r problem 6.54b}
(prob <- pexp(180, rate = 1/120, lower.tail = FALSE))
```
The probability that the watch will not have to be reset in at least 180 days is
$p=`r round(prob,4)`$.

# Problem 6.56
The number of bad cheques that a bank receives during a 5-hour business day is a
Poisson random variable with $\lambda=2$. What is the probability that it will 
not receive a bad cheque on any one day during the first two hours of business?

We need to scale our rate, because $\lambda$ is describing the rate of bad
cheques received over a 5 hour period, but we are dealing with only the first 2
hours of the business day.

```{r problem 6.56}
(prob <- ppois(0, 4/5))
```
The probability that the bank will not receive any bad cheques in the first two
hours of business on any given day is $p=`r round(prob, 4)`$.

# Problem 6.58
If the annual proportion of erroneous income tax returns filed with the IRS can
be looked upon as a random variable having a beta distribution with $\alpha=2$
and $\beta=9$, what is the probability that in any given year there will be 
fewer than 10% erroneous returns?

```{r problem 6.58}
(prob <- pbeta(0.1, shape1=2, shape2=9))
```
The probability that the IRS will receive fewer than 10% erroneous returns is
$p=`r round(prob,4)`$.

The annual expected proportion of erroneous income tax returns is given by
$$
E[X] = \frac{\alpha}{\alpha+\beta} = \frac{2}{2+9} = \frac{2}{11} \approx0.22.
$$

# Problem 6.59
A certain kind of appliance requires repairs on the average once every 2 years.
Assuming that the times between repairs are exponentially distributed, what is 
the probability that such an appliance will work at least 3 years without
requiring repairs?

```{r problem 6.59}
(prob <- pexp(3, rate = 1/2, lower.tail = FALSE))
```
The probability that an appliance of this kind going at least 3 years without
requiring repairs is $p=`r round(prob, 4)`$.

# Problem 6.61
Suppose that the service life in hours of a semiconductor is a random variable 
having a Weibull distribution with $\alpha=0.025$ and $\beta=0.500$.  
a) How long can such a semiconductor be expected to last?

```{r problem 6.61a}
qweibull(0.5, shape = 0.5, scale = 1/(0.025)^2)
```
A semiconductor can be expected to last about 768.7 hours using these 
distribution parameters.

b) What is the probability that such a semiconductor will still be in operating
condition after 4000 hours?

```{r problem 6.61b}
(prob <- pweibull(4000, shape = 0.5, scale = 1/(0.025)^2, lower.tail = FALSE))
```
The probability that one of these semiconductors continuing to operate beyond
4000 hours is $p=`r round(prob, 4)`$.

# Problem 6.70
Suppose that during periods of meditation the reduction of a person's oxygen
consumption is a random variable having a normal distribution with $\mu=37.6$ cc
per minute and $\sigma=4.6$ cc per minute. Find the probabilities that during a
period of meditation a person's oxygen consumption will be reduced by  
a) at least 44.5 cc per minute

```{r problem 6.70a}
mu <- 37.6
sigma <- 4.6

(prob <- pnorm(44.5, mean=mu, sd=sigma, lower.tail = FALSE))
```
The probability that during a period of meditation a person's oxygen consumption
will be reduced by at least 44.5 cc per minute is $p=`r round(prob, 4)`$.

b) at most 35.0 cc per minute

```{r problem 6.70b}
(prob <- pnorm(35.0, mean=mu, sd=sigma))
```
The probability that during a period of meditation a person's oxygen consumption
will be reduced by at most 35.0 cc per minute is $p=`r round(prob, 4)`$.

c) anywhere from 30.0 to 40.0 cc per minute

```{r problem 6.70c}
probs <- pnorm(c(30.0,40.0), mean=mu, sd=sigma)
(diff_prob <- probs[2] - probs[1])
```

The probability that during a period of meditation a person's oxygen consumption
will be reduced by between 30.0 and 40.0 cc per minute is $p=`r round(diff_prob, 4)`$.

